\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{isabelle,isabellesym}
\usepackage{amsmath,amssymb}
\usepackage{pdfsetup}

\urlstyle{rm}
\isabellestyle{it}

\begin{document}

\title{Verified Compartmental Models: A Reusable Framework\\with SIR Instantiation}
\author{David D.}
\maketitle

\begin{abstract}
We present a reusable Isabelle/HOL framework for certified reasoning about
compartmental ODE models, together with a complete instantiation for the
classical SIR (Susceptible--Infectious--Recovered) epidemic model.

The framework's central result is a generic forward invariance theorem for
linear ODEs: any equation $X' = f(t) \cdot X$ with continuous $f$ preserves
the sign of initial conditions, proved via the integrating factor method
without requiring Picard--Lindel\"of existence/uniqueness.

The SIR instantiation proves: forward invariance of the nonnegative orthant
from initial conditions alone; explicit solution formulas; conservation of
total population; monotonicity of $S$ and $R$; the Kermack--McKendrick
phase-plane invariant ($I + S - (\gamma/\beta)\ln S = \text{const}$);
the basic reproduction number $R_0 = \beta S(a)/\gamma$ with threshold
conditions; global monotone decline of $I$ when $R_0 \le 1$; and boundedness.

All results are fully proved with zero \texttt{sorry}/\texttt{admit}/\texttt{oops}.
\end{abstract}

\tableofcontents

\section{Introduction}

Compartmental models are a cornerstone of mathematical epidemiology,
pharmacokinetics, and systems biology. The SIR model, introduced by
Kermack and McKendrick~\cite{kermack1927}, partitions a population into susceptible~(S),
infectious~(I), and recovered~(R) compartments governed by a system of ODEs.

This development formalizes the key qualitative properties of the SIR model
in Isabelle/HOL, structured as a reusable compartmental model framework
with the SIR model as its first instantiation.

\paragraph{Contributions.}
\begin{itemize}
\item A generic linear ODE solution theorem via the integrating factor,
      yielding scalar sign preservation for arbitrary continuous coefficients.
\item Forward invariance: nonneg initial conditions imply nonneg trajectories,
      proved without Picard--Lindel\"of.
\item A generic three-compartment conservation theorem.
\item Monotonicity corollaries for compartments with signed derivatives.
\item A locale-based SIR formalization proving forward invariance, explicit
      solutions, conservation, monotonicity, the Kermack--McKendrick invariant,
      the basic reproduction number $R_0$ with global epidemic-decline theorem,
      and the stationary infection condition.
\end{itemize}

% generated text of all theories
\input{session}

\bibliographystyle{abbrv}
\bibliography{root}

\end{document}
